---
title: "mlr3learners.lgbpy: Mulitclass Classification GPU Example"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    keep_md: true
vignette: >
  %\VignetteIndexEntry{mlr3learners_lgbpy_multiclass_gpu}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

```{r setup}
library(mlr3)
library(mlr3learners.lgbpy)
library(lightgbm.py)
```

# Install the package

In order to use the `mlr3learners.lgbpy` R package, please make sure, the reticulate package is configured properly on your system (reticulate version >= 1.14) and is pointing to a python environment. If not, you can e.g. install `miniconda`:

```{r eval=FALSE}
reticulate::install_miniconda(
  path = reticulate::miniconda_path(),
  update = TRUE,
  force = FALSE
)
```
```{r}
reticulate::py_config()
```

Before you can use the `mlr3learners.lgbpy` package with GPU acceleration, you need to compile the `lightgbm` python package according to [its documentation](https://github.com/microsoft/LightGBM/blob/master/python-package/README.md). 
You can compile the GPU version on Linux/ in docker with the following commands: 

```{bash}
pip install lightgbm --install-option=--gpu
```

# Load the mlr3 task

```{r}
task <- mlr3::tsk("iris")
target_col <- "Species"
id_col <- NULL
```

To evaluate the model performance, the dataset is split into a training set and a test set with `sklearn_train_test_split`. This function is a wrapper around python sklearn's [sklearn.model_selection.train_test_split](https://scikit-learn.org/stable/modules/generated/sklearn.model_selection.train_test_split.html) method and ensures a stratified sampling for categorical target variables. 

```{r}
split <- lightgbm.py::sklearn_train_test_split(
  task$data(),
  target_col,
  split = 0.7,
  seed = 17,
  return_only_index = TRUE,
  stratify = TRUE
)
table(task$data()[split$train_index, task$target_names, with = F])
table(task$data()[split$test_index, task$target_names, with = F])
```

# Instantiate the lightgbm learner 

Initially, the `classif.lgbpy` class needs to be instantiated: 

```{r}
learner <- mlr3::lrn("classif.lgbpy")
```

# Configure the learner 

In the next step, some parameters need to be set. `validation_split` can be set in order to further split the training data and evaluate the model performance during training against the validation set. The allowed value range is 0 < validation_split <= 1. This parameter can also be set to "1", taking the whole training data for validation during the model training. For reproducibility, please use the `split_seed` argument.  
Almost all possible parameters have been implemented here. You can inspect them using the following command: 

```{r eval=FALSE}
learner$param_set
```

Also the parameters `num_boost_round` and `early_stopping_rounds` can be set here. Please refer to the [LightGBM manual](https://lightgbm.readthedocs.io) for further details these parameters.

```{r}
learner$id_col <- id_col
learner$early_stopping_rounds <- 10
learner$num_boost_round <- 100
```

In order to use the GPU acceleration, the parameter `device_type = "gpu"` (default: "cpu") needs to be set. According to the [LightGBM parameter manual](https://lightgbm.readthedocs.io/en/latest/Parameters.html), 'it is recommended to use the smaller `max_bin` (e.g. 63) to get the better speed up'. 

```{r}
learner$param_set$values <- list(
  "learning_rate" = 0.1,
  "seed" = 17L,
  "metric" = "multi_logloss",
  "device_type" = "gpu",
  "max_bin" = 63L
)
```

# Train the learner 

The learner is now ready to be trained by using its `train` function. 

```{r results='hide', message=FALSE, warning=FALSE, error=FALSE}
learner$train(task, row_ids = split$train_index)
```

# Evaluate the model performance 

The learner's `predict` function returns an object of mlr3's class `PredictionClassif`.

```{r}
predictions <- learner$predict(task, row_ids = split$test_index)
head(predictions$response)
```

It includes also a confusion  matrix:

```{r}
predictions$confusion
```

Further metrics can be calculated by using mlr3 measures:

```{r}
predictions$score(mlr3::msr("classif.logloss"))
```

The variable importance plot can be calculated by using the learner's `importance` function:

```{r}
importance <- learner$importance()
importance
```
