---
title: "mlr3learners.lightgbm: Hyperparameter Tuning Example"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    keep_md: true
vignette: >
  %\VignetteIndexEntry{mlr3learners_lightgbm_tuning}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
dir.create("./png")
```

```{r setup}
library(mlr3)
library(mlr3learners.lightgbm)
library(mlbench)
library(lightgbm.py)
library(paradox)
```

```{r include=FALSE}
reticulate::py_config()
lightgbm.py::install_py_lightgbm()
```

# Load the dataset

The data must be provided as a `data.table` object. To simplify the subsequent steps, the target column name and the ID column name are assigned to the variables `target_col` and `id_col`, respectively. 

```{r}
data("PimaIndiansDiabetes2")
dataset <- data.table::as.data.table(PimaIndiansDiabetes2)
target_col <- "diabetes"
id_col <- NULL
```

# Define an mlr3 task 

```{r}
task <- TaskClassif$new(
  id = "PimaIndiansDiabetes2",
  target = target_col,
  backend = dataset
)
```

# Instantiate the lightgbm learner 

Initially, the `classif.lightgbm` class needs to be instantiated: 

```{r}
learner <- mlr3::lrn("classif.lightgbm")
```

# Configure the learner 

```{r}
learner$split_seed <- 2
learner$validation_split <- 0.7
learner$early_stopping_rounds <- 1000
learner$num_boost_round <- 5000
```

Next, the learner parameters need to be set. At least, the `objective` parameter needs to be provided! Almost all possible parameters have been implemented here. You can inspect them using the following command: 

```{r eval=FALSE}
View(learner$param_set$default)
```

Please refer to the [LightGBM manual](https://lightgbm.readthedocs.io) for further details on these parameters. 

```{r}
learner$param_set$values <- list(
  "objective" = "binary",
  "learning_rate" = 0.01,
  "seed" = 17L,
  "bagging_freq" = 5L
)

tune_ps <- ParamSet$new(list(
  ParamDbl$new("bagging_fraction", lower = 0.4, upper = 1),
  ParamInt$new("min_sum_hessian_in_leaf", lower = 1e-3, upper = 5)
))
```

# Resampling strategy

```{r}
resamp <- mlr3::rsmp("cv")
measure <- mlr3::msr("classif.auc")
```

# Train the learner 

```{r}
terminate <- mlr3tuning::term("stagnation")

instance <- mlr3tuning::TuningInstance$new(
  task = task,
  learner = learner,
  resampling = resamp,
  measures = measure,
  param_set = tune_ps,
  terminator = terminate
)
instance
```

```{r eval=FALSE}
tuner <- mlr3tuning::tnr("grid_search", resolution = 5)
result <- tuner$tune(instance)
```
